The GitHub Open Source Development Process \author{Kevin Peterson\\ Department of Information Technology\\ Clinical Informatics Support Systems\\ Mayo Clinic\\ 200 1st Street\\ Rochester, MN 55905\\ \small \texttt{peterson.kevin@mayo.edu} }

### 2013-03-11 paper/main.tex
Expanded the CTS2 part.

\subsection{Hypothesis}
\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsicly motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository. As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correllation between the two.\\

\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}. In the context of Git and GitHub, however, a \emph{fork} is a positive occurence for a project, as it signifies greater project involvement. As Git allows for easy merging of \emph{forks}, code contributions in the form of defect fixes can be incorporated quickly. Because of this, it is expected that \emph{Issue Resolution Time} will decrease as the number of \emph{forks} increases.\\

@@

This is, ultimetely, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.

\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}. The Apache Server project is an excellent case study in this type of Developer Hierarchy. Apache is built around a small set of {\it Core Developers}, followed by {\it Defect Repairers} and {\it Defect Reporters}. Each level of this hierarchy brings with it an order of magnitude increase in number of participants. The 10-15 {\it Core Developers} contribute around 80\% of the new functionality, while the rest of the 400 code contributors focus primarily on bug fixes. The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.\\

@@

\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\

\emph{Research Question 2: How is GitHub changing the OSS process?}\\
The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites. What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\

\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The Object Managment Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards. Potential standard summission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.

@@

The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis. Results were stored in a MySQL relational database management system.

@@

The database schema to store the data is decribed in \ref{fig:er_diagram}.

@@

\subsection{Hypotheses}
\emph{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed betwee the number of Watchers and Forks of a repository. This suggests that a correlation between Watchers and Forks exists.
\begin{figure}
\includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
\caption{Repostitory watchers and forks}
\end{figure}
The data reinforces the hypothesis, as does other research in the area. Even though there are aguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related. Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}. If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking). This could possibly contribute to the correlation of the two variables.

\emph{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for serveral reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests. TODOs, or general support questions. \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
\caption{Repostitory issue close time and number of forks}
\end{figure}

\emph{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_reporters_histogram.png}
\caption{Repostitory issue reporters}
\end{figure}
\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_histogram.png}
\caption{Repostitory committers}
\end{figure}
\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
\caption{Repostitory issue reporters and committers}
\end{figure}

\subsection{Research Questions}
\emph{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Individual committer commit percentage per repository}
\end{figure}

\emph{Research Question 2: How is GitHub changing the OSS process?}\\
Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook. In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.

\emph{Research Question 3: Can GitHub be used for more that code artifacts?}\\
As part of the OMG\textregistered standardization process, the CTS2 Specification used GitHub issue tracker to track specification changes.

>>>>

\subsection{Hypothesis}
\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
In Open Source Software, a feeling of community belonging can be intrinsicly motivating to developers\cite{lakhani2003hackers}. This feeling of belonging can be expressed passively, as a \emph{watch}, or actively, as a \emph{fork} of a repository. As \emph{watches} don't necessarily signal intent to contribute, a positive direction for an OSS project is to not only increase the number of \emph{watchers}, but also the number of \emph{forks}. As both signify a general general interest in the project, it is expected that there be some correllation between the two.\\

\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
In OSS, a project \emph{fork} has at times carried negative connotations, and has even been referred to as a hazzard\cite{kogut2001open}. In the context of Git and GitHub, however, a \emph{fork} is a positive occurence for a project, as it signifies greater project involvement. As Git allows for easy merging of \emph{forks}, code contributions in the form of defect fixes can be incorporated quickly. Because of this, it is expected that \emph{Issue Resolution Time} will decrease as the number of \emph{forks} increases.\\

@@

This is, ultimetely, a proof of ``Linus Law"\cite{raymond1999cathedral} in the context of GitHub. In other words, the more exposure the code gets (in terms of forks), the easier bugs will be to find and fix.\\

\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude}\\
Research into the Apache Server project observed that there were far more \emph{Issue Reporters} than there were code \emph{Committers}\cite{mockus2000case}. The Apache Server project is an excellent case study in this type of Developer Hierarchy. Apache is built around a small set of {\it Core Developers}, followed by {\it Defect Repairers} and {\it Defect Reporters}. Each level of this hierarchy brings with it an order of magnitude increase in number of participants. The 10-15 {\it Core Developers} contribute around 80\% of the new functionality, while the rest of the 400 code contributors focus primarily on bug fixes. The {\it Defect Reporters} were by far the largest group, with over 3000 individuals submitting bug reports.\\

@@

\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
A smal core of developers, however, seems to be a historial fact of OSS development\cite{mockus2000case,mockus2002two,krishnamurthy2002cave}. At the extreme of this, a study of projects on the software hosting site Sourceforge\footnote{http://sourceforge.net} found that a large percentage of OSS development is done by lone developers. Is a small core of developers intrinsic to OSS development, or have the social aspects of GitHub and the distributed nature of Git changes this?\\

\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
The socal aspects of GitHub are an important part of the experience\cite{dabbish2012social}. User interactions and social pressures can drive OSS development in interesting ways. Social aspects of OSS development have existed before GitHub, with mailing lists\cite{mockus2000case}, gift giving mechanisms\cite{bergquist2008power}, and other code hosting sites. What is GitHub doing to facilitate change on OSS development, and how does it see OSS development changing moving forward?\\

\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The Object Managment Group\textregistered\footnote{http://www.omg.org} computer industry consortium focused on technology standards. Potential standard summission teams must go through a vetting process\cite{kobryn1999uml} which allows industry representatives to provide feedback on the standard. Gathering and recording this feedback is a challenge, but GitHub's issue tracking system may be able to streamline the process.

@@

The \textsc{store} function accepts as an input a given random repository. The purpose of this function is to persist the given repository to a database for further analysis. Results were stored in a MySQL relational database management system. The database schema to store the data is decribed in figure \ref{fig:er_diagram}.

@@

@@

\subsection{Hypotheses}
\textbf{Hypothesis 1: As the number of Watchers increase, the number of Repository Forks will increase.}\\
A correlation coefficient of \input{watcher_forks_scatterplot-correlation} was observed betwee the number of Watchers and Forks of a repository. This suggests that a correlation between Watchers and Forks exists.
\begin{figure}
\includegraphics[height=3in,width=3in]{images/watcher_forks_scatterplot.png}
\caption{Repostitory watchers and forks}
\end{figure}
The data reinforces the hypothesis, as does other research in the area. Even though there are aguments that the number of Forks is the true measure of project success\cite{baudry2012towards}, the data suggests that the two measures are related. Also, \textit{Dabbish et al.} have noted that the number of Watches can be a key for others to gauge interest in a project, and thus determine if they should participate \cite{dabbish2013leveraging}. If the number of Watchers of a repository serves as a social cue for participation, even ``passive'' involvement with a project (Watching) could influence the ``active'' participation (Forking). This could possibly contribute to the correlation of the two variables.

\textbf{Hypothesis 2: As the number Repository Forks increase, the Issue Resolution Time will decrease.}\\
No significant correlation was found between the number of repository forks and issue resolution time. It is important to note, however, that issue resolution time is not necessarily related to software product quality in the context of GitHub. We can assert that for serveral reasons. \textit{First}, GitHub issues are not necessarily product defects. Issues may be feature requests. TODOs, or general support questions. \textit{second}, GitHub projects may have primary issue tracking services elsewhere, while using the GitHub issues as secondary services or not at all. \textit{Third}, Pham \textit{et al.} explore an emerging ``testing culture'' in GitHub -- something which could have sizable impacts on software quality but wouldn't necessarily be reflected in this metric \cite{phamcreating}.
\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_close_time_forks_scatterplot.png}
\caption{Repostitory issue close time and number of forks}
\end{figure}

\textbf{Hypothesis 3: There will be more Issue Reporters than Committers by an order of magnitude.}\\
\begin{figure}
\includegraphics[height=3in,width=3in]{images/issue_reporters_committers_scatterplot.png}
\caption{Repostitory issue reporters and committers}
\label{fig:issue_reporters_committers_scatterplot}
\end{figure}
Data does not support an exponentially increasing number of distinct Issue Reporters as compared to distinct Committers. Figure\ref{fig:issue_reporters_committers_scatterplot}, does however suggest that that the two variables may be related. The hypothesis implies that there will be a correlation between the two variables -- which is demonstrated by a reasonably strong correlation cofficient of \input{issue_reporters_committers_scatterplot-correlation}. This does not seem to align with findings in the Apache Server project\cite{mockus2000case}. More investigation would be necessary to further examine this finding.

\subsection{Research Questions}
\textbf{Research Question 1: Are GitHub projects primarily focused around a small set of core Committers?}\\
\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_percentage_pie_chart.png}
\caption{Individual committer commit percentage per repository}
\label{fig:committers_percentage_pie_chart}
\end{figure}
The data in figure \ref{fig:committers_percentage_pie_chart} represents commit percentage breakdown. The inner percentages represent the percentage of committers in a project that that have committed the outer percentage amount of commits to that project, or

 \[ P = \left\{ \frac{ C_u } { |C| } \times 100 \Big| u \in U \right\} \]

where, given the set of commits of a repository $C$ and the set of committers to the repository $U$, the commits by each user $C_u$ are divided by the total number of commits $|C|$, yielding the set of contribution percentages $P$ for the repository. Elements of $P$ are then assigned to ranges $0-5\%$, $5-10\%$, $10-50\%$, $50-95\%$, and $95-100\%$ (with shared values inclusive to the larger range). The distribution of this range assignments is plotted on figure \ref{fig:committers_percentage_pie_chart}.

There does appear to be evidence of a small set of core developers, denoted by the relatively high percentage of developers that have authored $95-100\%$ of the commits of a given repository. Most notably, however, is that the data shows that by a large percentage of developers committed small amounts ($0-5\%$ of the total repository commits). This means that a small core are not entirely monopolizing the commits, but rather opportunity exists for users to contribute, even if in small ways.

\begin{figure}
\includegraphics[height=3in,width=3in]{images/committers_histogram.png}
\caption{Count of individual committers per repository}
\label{fig:committers_histogram}
\end{figure}

When looking strictly at the number of committers to a repository, however, the notion of a small set of core Committers is strongly supported, as the vast majority of repositoris have fewer than 10 committers (figure \ref{fig:committers_histogram}).

\textbf{Research Question 2: How is GitHub changing the OSS process?}\\
Thung \textit{et al.} show that the social coding aspects of GitHub -- specifically the developer-to-developer connectivity -- enable high collaboration rates.\cite{thung2013network} In fact, the developer connectivity was shown to be higher than SourceForge and even Facebook. In a recent interview with GitHub staff, the social aspects of GitHub are referred to as ``unique and powerful\cite{begel2013social},'' which seems to suggest that this is a strong underlying product goal for.

\textbf{Research Question 3: Can GitHub be used for more that code artifacts?}\\
The OMG\textregistered standardization process is a process by which products strive to produce an interoperability statndard by way of an industry standards consortium.\cite{omgTechProcess}
The standardization process itself aims to, at its conclusion, produce a set of model artificats containing sufficient detail such that when interpreted and implemented, produce interoperable software.
At a conceptual level, this can be thought of as a software requirements engineering process as much as a standardization process. If so,  of the five main tasks of requirements engineering, Elicitation, Analysis and Negotiation, Documentation, Validation, and Management,\cite{sommerville1998requirements} we can show that GitHub is an apporiate forum for many of them.

As a case study, the CTS2 OMG\textregistered  Specification\footnote{https://github.com/cts2/cts2-specification} used GitHub issue tracker to track specification changes. Elicitation, Analysis and Negotiation were natural fits for the issue tracker. This was chosen in part because the barrier to participation was very small (only a GitHub account was needed), and issues tracker itself allowed for dialogs in the form of comments. Because customer involvement is critical to these activities\cite{paetsch2003requirements}, ease of contribution was a main concern. OMG has a predescribed format for requirements Documentation, but using the GitHub issue API\footnote{http://developer.github.com/v3/issues/}, much of this documentation can be automated.
Validation and Management are areas of requirements engineering that GitHub has interesting solutions for. First, in order to track (or validate) changes to the specification, when the actual specification is modified, the commit that modified it can be traced to a named issue by simply putting the issue number in the commit log. This level of traceability is powerful, as it allows every change to the specification artifacts to be tracked consistently to an open issue. As the normal OMG process proceeds, requirements can be managed through ``tagging'' the issues. By allowing issues to be grouped by arbitrary tags, users are more free to adapt their own workflow. For instance, when community voting on issues needed to be done for CTS2, a simple ``ready for vote'' tag signified to the voting board exactly what was under consideration.

As mentioned, the CTS2 specification made heavy use of the GitHub issue tracker, as well as another OMG specification, ServD\footnote{https://github.com/servd/servd-specification}. For open source specifications, or requirements engineering activities, GitHub has shown to be a viable platform.

\section{Conclusion}
